Abstract Binding Trees

### 2015-11-05 paper.tex
Remove code from document for now until it settles down
formatting of metavariables [try using fraktur]

% make lualatex play nice with agda
\makeatletter
\ifluatex
   \newwrite\CATCHPATCHOUT
\fi
\def\CatchFBT@Fin@l#1[#2]{%
   \begingroup%
      \endlinechar\m@ne \makeatletter #2%
      \ifluatex%
         \immediate\openout\CATCHPATCHOUT=\jobname.catchpatch%
         \immediate\write\CATCHPATCHOUT {\CatchFBT@tok{\the\CatchFBT@tok}}%
         \immediate\closeout\CATCHPATCHOUT%
         \@@input \jobname.catchpatch%
      \else%
      \scantokens\expandafter{%
         \expandafter\CatchFBT@tok\expandafter{\the\CatchFBT@tok}}%
      \fi%
      \CatchFBT@IsAToken{#1}%
         {\global#1\expandafter{\the\CatchFBT@tok}}%
         {\xdef#1{\the\CatchFBT@tok}}%
      \ifx\CatchFBT@tok#1\else\global\CatchFBT@tok{}\fi%
   \endgroup%
}% \CatchFBT@Final
\makeatother

@@

\newcommand\MkSet[1]{\left\{#1\right\}}
\newcommand\Member[2]{\IMode{#1}\in\IMode{#2}}
\newcommand\EqMember[3]{\IMode{#1}=\IMode{#2}\in\IMode{#3}}
\newcommand\Of[2]{\IMode{#1} : \IMode{#2}}

@@

\newcommand\IsOperator[3]{\IMode{#1}\Vdash \IMode{#2} : \OMode{#3}}

@@

\newcommand\Grothendieck[2]{\oint^{#1} #2#1}
\newcommand\Pair[2]{\left\langle #1, #2 \right\rangle}
\newcommand\Tuple[1]{\left\langle #1 \right\rangle}
\newcommand\FLift[2]{
  \Operators\langle#1%
  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle%
}
\newcommand\PairTimes[2]{\langle #1\, {\scriptstyle {{\times}}}\, #2 \rangle}
\newcommand\SSingleton{\bullet}
\newcommand\Forget[1]{|\, #1\, |}

\newcommand\Comma[2]{#1 \downarrow #2}
\newcommand\Coslice[2]{#1\, /\, #2}

@@

\newcommand\MV[1]{\textsc{#1}}
\newcommand\MApp[3]{#1\{#2\}(#3)}
\newcommand\App[2]{#1(#2)}
\newcommand\Dom[1]{\left\vert #1\right\vert}
\newcommand\FV[1]{\mathbf{FV}\left(#1\right)}
\newcommand\FS[1]{\mathbf{FS}\left(#1\right)}
\newcommand\NotIn[2]{\IMode{#1}\notin\IMode{#2}}
\newcommand\SortExp{\mathsf{exp}}

@@

\newcommand\Rename[2]{%
  #2\upharpoonright #1%
}
\newcommand\IsRenaming[3]{%
  \Rename{\IMode{#1}}{\IMode{#2}}%
  \leadsto\OMode{#3}%
}

@@

\newcommand\SemBrackets[1]{\left\llbracket #1\right\rrbracket}
\newcommand\Yoneda[1]{\boldsymbol{y}(#1)}

@@

\[
  \infer{
    \IsArity{\MkArity{v_0,...,v_n}{s}}
  }{
    \IsSort{s} &
    \HypJ{\IsValence{v_i}}{\Leq{i}{n}}
  }
\]

@@

Using this, we can define a tensor $P\odot Q$ for
$\Of{P,Q}{\widehat{\HypCat}^\Sorts}$ as follows:
\[
  \HypJ{
    \Define{(P\odot Q)_s}{
      P_s\bullet Q
    }
    \quad
  }{
    \Member{s}{\Sorts}
  }
\]

@@

\begin{remark}
  A presheaf $P$ with a substitution monoidal structure $\Pair{\nu}{\varsigma}$
  also forms a \emph{relative monad}\footnote{Relative monads are skew-monoids
    in the skew-monoidal structure induced by a functor category
    $\mathbf{C}^\mathbf{J}$ \cite{altenkirch-chapman-uustalu:2015}. However,
    relative monads can also be understood without reference to monoids by
    viewing them as a generalization of \emph{extension systems} or
    \emph{Kleisli structures}. This latter presentation of monads, popular in
    computer science and logic, does not involve iteration of the monad
    functor.} on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
  Consequently, the monoid multiplication $\varsigma$ can be replaced with an
  extension operation:
  \[
    \begin{tikzcd}[ampersand replacement = \&, sep = huge]
      {\IMode{P(\,=\,)^{V(\,-\,)}}} \arrow[r, "(\,-\,)^{\circledast}"]
      \& {\IMode{P(\,=\,)^{P(\,-\,)}}}
    \end{tikzcd}
  \]
\end{remark}

@@

For each signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we have an
endofunctor
$\Of{\SigFun{\Sigma}}{\widehat{\HypCat}^\Sorts\to\widehat{\HypCat}^\Sorts}$,
which is defined as follows:
\begin{align*}
  \IMode{\SigFun{\Sigma}(X)_s}
    &\triangleq
    \OMode{
      \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{s}}}
      \prod_{\MkValence{\vec{p}}{\vec{q}}{s'}\in\vec{v}}
      X_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}}
    }
\end{align*}

@@

The metavariable, symbol and variable contexts are interpreted in a model $P$ as an
environment presheaf in the following way:
\[
  \Define{
    \SemBrackets{
      \Omega\triangleright\Upsilon\parallel\Gamma
    }_P
  }{
    \left(\prod_{(\MV{m}:\MkValence{\vec{p}}{\vec{q}}{s})\in\Omega}
    P_{s}^{\Yoneda{\vec{p}\parallel\vec{q}})}\right)
    \times
    S^{\{\Upsilon\}}
    \times
    V^{[\Gamma]}
  }
\]

@@

Crary has demonstrated a first-order encoding of contexts in the logical
framework in bijection with actual LF-contexts \cite{crary:2009}, which has
been successfully used in large-scale mechanization efforts, including that of
Standard ML \cite{lee-crary-harper:2007} and the Edinburgh Logical Framework
itself \cite{martens-crary:2012}.

@@

\section{Code}

\CatchFileBetweenTags\tmp{Code.tex}{agda}

\subsection{Kan Extensions}

\subsubsection{Lan}

\ExecuteMetaData[Code.tex]{lang}\\

\ExecuteMetaData[Code.tex]{lan}

\subsubsection{Ran}

\ExecuteMetaData[Code.tex]{rang}\\

\ExecuteMetaData[Code.tex]{ran}

\subsection{Variables and Symbols}

\ExecuteMetaData[Code.tex]{var}\\

\ExecuteMetaData[Code.tex]{sym}

\subsection{Contexts}

\ExecuteMetaData[Code.tex]{ctx}\\

\ExecuteMetaData[Code.tex]{sctx}\\

\ExecuteMetaData[Code.tex]{ctxdom}\\

\ExecuteMetaData[Code.tex]{ctxidx}\\

\ExecuteMetaData[Code.tex]{elem}

\subsection{Signatures}

\ExecuteMetaData[Code.tex]{sign}

\subsection{Trees}

\ExecuteMetaData[Code.tex]{trees}

\subsection{Substitution Monoidal Structure}

\ExecuteMetaData[Code.tex]{H}\\

\ExecuteMetaData[Code.tex]{HHat}\\

\ExecuteMetaData[Code.tex]{V}\\

\ExecuteMetaData[Code.tex]{tensor0}\\

\ExecuteMetaData[Code.tex]{tensor1}\\

\ExecuteMetaData[Code.tex]{extension}

>>>>

% make lualatex play nice with agda
\makeatletter
\ifluatex%
   \newwrite\CATCHPATCHOUT%
\fi
\def\CatchFBT@Fin@l#1[#2]{%
   \begingroup%
      \endlinechar\m@ne\makeatletter #2%
      \ifluatex%
         \immediate\openout\CATCHPATCHOUT=\jobname.catchpatch%
         \immediate\write\CATCHPATCHOUT{\CatchFBT@tok{\the\CatchFBT@tok}}%
         \immediate\closeout\CATCHPATCHOUT%
         \@@input\jobname.catchpatch%
      \else%
      \scantokens\expandafter{%
         \expandafter\CatchFBT@tok\expandafter{\the\CatchFBT@tok}}%
      \fi%
      \CatchFBT@IsAToken{#1}%
         {\global#1\expandafter{\the\CatchFBT@tok}}%
         {\xdef#1{\the\CatchFBT@tok}}%
      \ifx\CatchFBT@tok#1\else\global\CatchFBT@tok{}\fi%
   \endgroup%
}% \CatchFBT@Final
\makeatother

@@

\newcommand\MkSet[1]{\left\{#1\right\}}
\newcommand\Member[2]{\IMode{#1}\in\IMode{#2}}
\newcommand\EqMember[3]{\IMode{#1}=\IMode{#2}\in\IMode{#3}}
\newcommand\Of[2]{\IMode{#1}: \IMode{#2}}

@@

\newcommand\IsOperator[3]{\IMode{#1}\Vdash\IMode{#2}: \OMode{#3}}

@@

\newcommand\Grothendieck[2]{\oint^{#1} #2#1}
\newcommand\Pair[2]{\left\langle#1, #2 \right\rangle}
\newcommand\Tuple[1]{\left\langle#1 \right\rangle}
\newcommand\FLift[2]{
  \Operators\langle#1%
  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle%
}
\newcommand\PairTimes[2]{\langle#1\, {\scriptstyle{{\times}}}\, #2 \rangle}
\newcommand\SSingleton{\bullet}
\newcommand\Forget[1]{|\, #1\, |}

\newcommand\Comma[2]{#1 \downarrow#2}
\newcommand\Coslice[2]{#1\, /\, #2}

@@

\newcommand\MV[1]{\textsc{#1}}
\newcommand\MApp[3]{#1\{#2\}(#3)}
\newcommand\App[2]{#1(#2)}
\newcommand\Dom[1]{\left\vert#1\right\vert}
\newcommand\FV[1]{\mathbf{FV}\left(#1\right)}
\newcommand\FS[1]{\mathbf{FS}\left(#1\right)}
\newcommand\NotIn[2]{\IMode{#1}\notin\IMode{#2}}
\newcommand\SortExp{\mathsf{exp}}

@@

\newcommand\Rename[2]{%
  #2\upharpoonright#1%
}

\newcommand\IsRenaming[3]{%
  \Rename{\IMode{#1}}{\IMode{#2}}%
  \leadsto\OMode{#3}%
}

@@

\newcommand\SemBrackets[1]{\left\llbracket#1\right\rrbracket}
\newcommand\Yoneda[1]{\boldsymbol{y} (#1)}

@@

\[
  \infer{
    \IsArity{\MkArity{v_0,\ldots,v_n}{s}}
  }{
    \IsSort{s} &
    \HypJ{\IsValence{v_i}}{\Leq{i}{n}}
  }
\]

@@

Using this, we can define a tensor $P\odot Q$ for
$\Of{P,Q}{\widehat{\HypCat}^\Sorts}$ as follows:
\[
  \HypJ{
    \Define{{(P\odot Q)}_{s}}{
      P_s\bullet Q
    }
    \quad
  }{
    \Member{s}{\Sorts}
  }
\]

@@

\begin{remark}
  A presheaf $P$ with a substitution monoidal structure $\Pair{\nu}{\varsigma}$
  also forms a \emph{relative monad}\footnote{Relative monads are skew-monoids
    in the skew-monoidal structure induced by a functor category
    $\mathbf{C}^\mathbf{J}$~\cite{altenkirch-chapman-uustalu:2015}. However,
    relative monads can also be understood without reference to monoids by
    viewing them as a generalization of \emph{extension systems} or
    \emph{Kleisli structures}. This latter presentation of monads, popular in
    computer science and logic, does not involve iteration of the monad
    functor.} on the functor $\Of{V}{\widehat{\HypCat}^\Sorts}$ of variables.
  Consequently, the monoid multiplication $\varsigma$ can be replaced with an
  extension operation:
  \[
    \begin{tikzcd}[ampersand replacement = \&, sep = huge]
      {\IMode{{P(\,=\,)}^{V(\,-\,)}}} \arrow[r, "{(\,-\,)}^{\circledast}"]
      \& {\IMode{{P(\,=\,)}^{P(\,-\,)}}}
    \end{tikzcd}
  \]
\end{remark}

@@

For each signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we have an
endofunctor
$\Of{\SigFun{\Sigma}}{\widehat{\HypCat}^\Sorts\to\widehat{\HypCat}^\Sorts}$,
which is defined as follows:
\begin{align*}
  \IMode{{\SigFun{\Sigma}(X)}_{s}}
    &\triangleq
    \OMode{
      \coprod_{\vartheta\in \Operators_{\MkArity{\vec{v}}{s}}}
      \prod_{\MkValence{\vec{p}}{\vec{q}}{s'}\in\vec{v}}
      X_{s'}^{\Yoneda{\vec{p}\parallel\vec{q}}}
    }
\end{align*}

@@

The metavariable, symbol and variable contexts are interpreted in a model $P$ as an
environment presheaf in the following way:
\[
  \Define{
    \SemBrackets{
      \Omega\triangleright\Upsilon\parallel\Gamma
    }_P
  }{
    \left(\prod_{(m:\MkValence{\vec{p}}{\vec{q}}{s})\in\Omega}P_{s}^{\Yoneda{\vec{p}\parallel\vec{q}}}\right)
    \times
    S^{\{\Upsilon\}}
    \times
    V^{[\Gamma]}
  }
\]

@@

Crary has demonstrated a first-order encoding of contexts in the logical
framework in bijection with actual LF-contexts~\cite{crary:2009}, which has
been successfully used in large-scale mechanization efforts, including that of
Standard ML~\cite{lee-crary-harper:2007} and the Edinburgh Logical Framework
itself~\cite{martens-crary:2012}.

@@
